---
title: "P8105_HW2_WZ2631"
author: "wz2631"
date: "2022-10-06"
output: html_document
---


```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

### Import and clean data firstly.
```{r}
trans_ent = 
  read_csv(
    "/Users/pap/Desktop/R_space/P8105_HW2_WZ2631/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

### The number of unique stations is 465.
```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

### The number of ADA compliant stations is 84.
```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

### The proportion of station entrances / exits without vending allow entrance is 0.3770492.
```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

### The numeber of stations that serve the A train is 60, of which 17 stations are ADA compliant.
```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

## Problem 2

### Import and clean data of Mr. Trash Wheel.
```{r}
df_mr = 
 read_excel(
   "/Users/pap/Desktop/R_space/P8105_HW2_WZ2631/Trash Wheel Collection Data.xlsx", 
   sheet = 1, 
   range = "A2:N549") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls), 
         data_source = "mr_trash_wheel",
         year = as.integer(year)) 
```

### Import and clean data of Professor. Trash Wheel.
```{r}
df_prof = 
 read_excel(
   "/Users/pap/Desktop/R_space/P8105_HW2_WZ2631/Trash Wheel Collection Data.xlsx", 
   sheet = 2, 
   range = "A2:M96") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(data_source = "prof_trash_wheel") 
```

### Combine the data of Mr. Trash Wheel and Professor. Trash Wheel.
```{r}
df_trash =
  bind_rows(df_mr, df_prof) %>%
  janitor::clean_names() 
df_trash
```

### Description and answer.
The time range of this observation is from May 2014 to July 2022. There are 15 variables and 641 observations, corresponding to the 15 columns and 641 rows of the "df_trash" dataframe respectively.
The key variables include date, weight_tons and volume_cubic_yards, and different types of trash are described as well. The additional variable "data_source" was added to keep track of which one of the two Trash Wheels that collected the trash.
```{r}
weight_prof =
  df_trash %>%
  filter(data_source == "prof_trash_wheel") 
  
sum(weight_prof$weight_tons)
```
Therefore, the total weight of trash collected by Professor Trash Wheel was 190.12 tons.
```{r}
sportsballs2020_mr =
  df_trash %>%
  filter(data_source == "mr_trash_wheel", year == 2020)
  
sum(sportsballs2020_mr$sports_balls)
```
As a result, the total number of sports balls collected by Mr. Trash Wheel in 2020 was 856.

## Problem 3
### Import and clean data of `pols-month.csv` firstly.
```{r}
library(magrittr)
data_pols = 
  read_csv("/Users/pap/Desktop/R_space/P8105_HW2_WZ2631/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day"), sep = "-") %>%
  mutate(
    month = as.integer(month),
    month = month.abb[month],
    year = as.integer(year),
    president = ifelse(prez_dem == 1, "dem", "gop")) %>% 
  dplyr::select(-day, -prez_gop, -prez_dem)
data_pols
```

### Import and clean data of `snp.csv`.
```{r}
data_snp = 
  read_csv("/Users/pap/Desktop/R_space/P8105_HW2_WZ2631/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), sep = "/") %>% 
  mutate(
    year = as.integer(year),
    year = ifelse(year > 2049, year - 100, year),
    month = as.integer(month),
    month = as.character(month),
    day = as.integer(day)) %>% 
    select(-day) %>% 
  relocate(year, month)
data_snp
```

### Import and clean data of `unemployment.csv`. Prepare the data of `unemployment.csv` for merge by pivot_longer().
```{r}
data_unemployment =
  read_csv("/Users/pap/Desktop/R_space/P8105_HW2_WZ2631/unemployment.csv") %>% 
  janitor::clean_names() %>%
  mutate(year = as.integer(year)) %>%
  pivot_longer(
    jan:dec, 
    names_to = "month", 
    values_to = "unemployment_rate")
data_unemployment 
```
### Merge three datasets.
```{r}
snp_pols = left_join(data_pols, data_snp, by = c("year","month")) 
snp_pols_unemployment = left_join(snp_pols, data_unemployment, by = c("year","month")) 
snp_pols_unemployment
```

### Explain what each dataset contained, and describe the resulting dataset.
1. The dataset of data_pols has 822 rows and 9 columns.The key variables included are year, month, statistics of the political groups of policians, also the political group to which the president belonged. It ranges from Jan 1947 to Jun 2015.
2. The dataset of data_snp has 787 rows and 3 columns. The key variables included are year, month, and the corresponding closing values of stock index. It ranges from Jan 1950 to Jul 2015.
3. The dataset of data_unemployment has 816 rows and 3 columns. The key variables included are year, month, and unemployment_rate. It ranges from Jan 1948 to Dec 2015.
4. The dataset of snp_pols_unemployment merged the three datasets together, with 822 rows and 11 columns. Its key variables includes year, month, statistics of the political groups of policians, the political group to which the president belonged, the corresponding closing values of stock index, and unemployment_rate. It ranges from Jan 1947 to Jun 2015, displaying the characteristics of the above three datasets are comprehensively, indicating the potential correlation between economy, politics and people's livelihood.